Measurement of a dwarf galaxy rotational velocity from its Lyman-alpha line emission

### 2016-02-03 paper/letter.tex
nuevos limits en D y M_D

\begin{document}
\pagestyle{empty}
\noindent
\textbf{Lyman-alpha emission reveals an unusual fast rotating dwarf galaxy}  
\\
\\
JEFR$^{1}$, MCRG$^1$, JNGC$^2$,
\\
\\
\scriptsize
{$^1$ Universided de los Andes
\\
$^2$ Arizona
\normalsize
\\
\\
\textbf{The Lyman-alpha emission line is a strong indicator of star formation.
Observations of this line are central to construct samples of the most
distant star forming galaxies, which in turn are useful for studies in
cosmology and galaxy evolution.  
The intepretation of Ly-$\alpha$ observations has to go together with
computational models for the radiative transfer of Lyman-alpha photons.
Recent theoretical work suggests that galaxy rotation has a measurable
impact on the shape of the Lyman-alpha line.  
Here we report that lyman-alpha observations of a blue dwarf compact galaxy (TOL
1214-277) in the local universe, which had previously evaded theoretical 
interpretation, are naturally explained by rotational effects.
We constrain the rotational velocity, viewing angle and total
neutral hydrogen mass from the Lyman-alpha observations.
These values are in broad agreement with other observational
constraints, although  they point towards a rather atypical physical
nature for the source.  
Our results present a new observational method to estimate the rotational
velocity of dwarf galaxies.
Considering the expected similarities between the local and the most
distant dwarf galaxies, we anticipate that the Lyman-alpha line could be
used to constraint the rotational state of the neutral gas in the
highest redshift galaxies to be detected with next generation infrared
spectroscopic facilities such as the James Webb Space Telescope.}

@@

The receeding velocity is $7785\pm 50$km s$^{-1}$, which translates
into a distance of $106.6$ Mpc (Hubble constant 73 Mpc km$^{-1}$
s$^{1}$)
The metallicity is $\sim Z_{\odot}/24$ \cite{Izotov04} as derived from optical
spectroscopy. The observed flux for the Lyman alpha line is $\sim
8.1\times 10^{-14}$ erg cm$^{-2}$ s$^{-1}$ \AA$^{-1}$ \cite{Thuan97}
and a Equivalent Width of $70$\AA, with the emission comming from a
region with approximate diameter ?. There is an upper limit for the
integrated flux of $<0.10$ Jy km s$^{-1}$, which translates into a
upper limit for the HI mass of $M<2.65\times 10^{8}$ M$_{\odot}$
\cite{pustilnikmartin07}.

@@

For an homogeneous sphere the HI optical depth can be written
as $\tau = \sigma_0 n s$, where $\sigma_0$ is the optical depth at the
line's center, $n$ is the number density and $s$ is the sistem's
size. 
This can be rewritten in terms of the gas' temperature $T$ and column
density $N_{H}$as $\tau = 3.31 \times 10^{-14} (10^{4}\mathrm{K}/T)^{1/2}
(N_{H}/\mathrm{atoms\ cm}^{-2})$.

@@

\begin{equation}
M_{T} = \frac{V^{2}D}{G} = 2.16\times10^{5}
\left(\frac{V}{\mathrm{km\ s}^{-1}}\right)^2\left(\frac{D}{\mathrm{kpc}}\right) M_{\odot}
\end{equation}
\end{document}

>>>>

\begin{document}
\pagestyle{empty}
\noindent
\textbf{Lyman-alpha emission reveals an unusual fastly rotating
  compact dwarf galaxy}  
\\
\\
JEFR$^{1}$, MCRG$^1$, JNGC$^2$,
\\
\\
\scriptsize
{$^1$ Universidad de los Andes
\\
$^2$ Arizona
\normalsize
\\
\\
\textbf{
Compact dwarf galaxies (CDGs) are puzzling systems. 
They form large quantities of stars in spite of their lack of
heavy chemical elements which are present in other nearby galaxies. 
The conditions in their interstellar medium resemble the expected
pristine environment of the first galaxies. 
Until these early galaxy generations are observationally
detected, CDGs are the best present day population to test our ideas
on primordial galaxy formation and evolution. 
Here we report on one of such CDGs, Tololo 1214-277, which presents
features in its Lyman$\alpha$ emission thad had evaded theoretical
interpretation so far. 
We show that these special features can be naturally explained by
rotational effects.  
We use the Lyman$\alpha$ observations available for Tololo 1214-277 to
constrain the rotational velocity and total neutral hydrogen optical
depth.  
We find that the neutral hydrogen region in Tololo 1214-277 should
have a rotational velocity of $V_{r}=300$ km s$^{-1}$ and an optical
depth of $\tau=10^7$.  Considering previous
observational upper limits on  the total neutral hydrogen mass and
expected number density, we find that the diameter for the Ly$\alpha$
emission region is in the range of $110$ pc $<D<$ $340$
pc. Correspondingly the total dynamical mass in that region is between
$2.1\times 10^{9}$M$_{\odot}$ $<M_D<$  $6.6\times 10^{9}$M$_{\odot}$.
We anticipate that the Lyman-alpha line could be
used to constraint the rotational state of the neutral gas in the
highest redshift galaxies to be detected with next generation infrared
spectroscopic facilities such as the James Webb Space Telescope.}

@@

The receeding velocity is $7785\pm 50$km s$^{-1}$, which translates
into a distance of $106.6$ Mpc (Hubble constant 73 Mpc km$^{-1}$
s$^{1}$)
The metallicity is $\sim Z_{\odot}/24$ \cite{Izotov04} as derived from optical
spectroscopy.

The observed flux for the Lyman alpha line is $\sim
8.1\times 10^{-14}$ erg cm$^{-2}$ s$^{-1}$ \cite{Thuan97}
and a Equivalent Width of $70$\AA and its H$\beta$ flux is 
$1.62\times 10^{-14}$ erg cm$^{-2}$ s$^{-1}$ cm${-2}$
\cite{Izotov04} which gives a Ly$\alpha$/H$\beta$ flux ratio of
4.9$\pm$0.1. 
Comparing this ratio with the theoretical expectation from case B
recombination of $23.3$ \cite{Hummer1987} one can estimate an escape
fraction of $20$\% for Ly$\alpha$ radiation.

The optical emission  comes from a   region with approximate diameter
?? \cite{Fricke01}.

@@

There is an upper limit for the  
integrated flux of $<0.10$ Jy km s$^{-1}$, which translates into a
upper limit for the HI mass of $M<2.65\times 10^{8}$ M$_{\odot}$
\cite{pustilnikmartin07}. 
From the optical depth of $10^7$ and the non-detection in the HI line,
we have an upper limit for the size where the Lya emission comes of
$D<0.34$kpc.

 For an homogeneous sphere the HI optical depth from its can be
 written as $\tau = \sigma_0 n D/2$, where $\sigma_0=5.898\times
10^{-14}$cm$^{-2}$ is the Lyman$\alpha$  optical depth at the 
line's center, $n$ is the number density and $D$ is the sphere's
diameter. 
From this we can impose additional constrains on $D$ from the tipical
values of the Hydrogen number density and our constrain on
$\tau=10^{7}$.  Using a range of $1<n/\mathrm{atoms/cm}^{-3} <
10^{-3}$. This gives us a range of $0.11 < D/\mathrm{kpc}<100$. 
Together from the total HI mass we have thus that the HI region should
have a diameter of $0.11 < D/\mathrm{kpc}<0.34$.

This can be rewritten in terms of the gas' temperature $T$ and column
density $N_{H}$as $\tau = 3.31 \times 10^{-14} (10^{4}\mathrm{K}/T)^{1/2}
(N_{H}/\mathrm{atoms\ cm}^{-2})$.

@@

\begin{equation}
M_{T} = \frac{V^{2}D}{G} = 2.16\times10^{5}
\left(\frac{V}{\mathrm{km\ s}^{-1}}\right)^2\left(\frac{D}{\mathrm{kpc}}\right) M_{\odot}
\end{equation}

From this limit and the rotational velocity of $300$ km/s and the
limit in the size $D$ we have limits of for the dynamical mass of
$2.1\times 10^{9}$M$_{\odot}$ $<M_D<$  $6.6\times 10^{9}$M$_{\odot}$.

which is at least $7$ to $25$ times larger than the H$I$ mass.

\end{document}

%% interpretation
